Response of mesophyll conductance to light availability allows shade leaves to lie in wait

### 2015-08-03 manuscript/manuscript_wtc3.Rmd
finalized first draft

---
title: "Near instantaneous response of mesophyll conductance to light availability allows shade leaves to lie in wait"
author: ""
date: ""
output:
  word_document:
    reference_docx: manuscriptstyle.docx
  pdf_document: default
  html_document:
    number_sections: yes
csl: new-phytologist.csl
bibliography: references.bib
---

@@

\
In this study we use *Eucalyptus tereticornis* Sm. trees, planted in climate controlled whole tree chambers with ambient and elevated temperature (+3&deg;C) treatments, to empirically evaluate the distribution of resources and leaf physiological behaviour of sun and shade leaves. Our hypothesese are as follows:

@@

\
Leaf level gas exchange was measured with a standard 6 cm^2^ leaf chamber using a portable gas exchange system (LI-6400, LI-COR, Lincoln, NE, USA).  This system was coupled with a tunable diode laser (TDL;TGA100,Campbell Scientific, Inc., Logan, UT, USA) for concurrent measurements of online C isotope discrimination. The CO~2~ in the leaf chamber was set at ambient atmospheric [CO~2~] (400 ppm) with a flow rate of 200 $\mu$mols s^-1^.  Two identical gas exchanges systems were run simultaneously, one in each of a randomly chosen WTC for each temperature treatment. Leaf temperatures were controlled at the current AT or ET chamber air temperature. PPFD in the chamber was set to match the individual light environment of each leaf type (explained above). Sun flecks were then simulated for shade leaves by increasing the chamber PPFD to the light environment of the full sun leaf in the same tree. The maximum sunfleck response was then recorded once shade leaves re-stablized in the leaf cuvette (ca. 25 min).

@@

where $\Delta$~o~ is the observed discrimation and  $\Delta$~i~, $\Delta$~gm~ , $\Delta$~e~ and $\Delta$~f~ are the contributions to fractionataion if C~i~ = C~c~, g~m~, respiration and photorespiration, respectively. The equations for each are as follows:

\
(5)
$$
\begin{aligned}
\Delta_i = &\frac{1}{1-t}a'+\frac{1}{1-t}((1+t)b-a')\frac{C_i}{C_a}
\end{aligned}
$$

@@

where C~s~ is the CO~2~ partial pressure at the leaf surface, a~b~ (2.9&permil;) is the fractionation from boundary layer diffusion and a is the fractionation due to diffusion in air (4.4&permil;) [@evans1986carbon]. C~a~ and C~i~ are the atmospheric and intercellular partial pressures and $\Gamma_{*}$ is the compensation point in the absence of mitochondrial respiration in the light (R~d~). In this experiment both $\Gamma_{*}$ and R~d~ were derived using a standard Arrhenius function with parameters for *Eucalyptus globulis* from [@crous2012light]. The ternary effect corrections (t) are desribed by:

@@

The CO~2~ diffusion from the intercellular airspace to the chloroplast, g~m~, is given by its relationship to the CO~2~ assimilation rate (A) by:

@@

where C~c~ is the chloroplast CO~2~ partial pressure. Once g~m~ was calculated C~c~ and the drawdown of CO~2~ from the intercellular arispaces to the site of carboxylation were then estimated using Equation 11. Examples of this approach to measure gas exchange and C isotope discrimination are presented in [@evans2013temperature]. The variation in $\Delta$~o~ between sun and shade leaves and the simulated sunfleck where then compared as a function of C~i~:C~a~.

@@

where the term Hp~w~g represents the gravitational potential assumed to be minimal in this experiment due to small tree size [@whitehead1998regulation]. Leaf level instantaneous transpiration efficiency (ITE) was calculated as leaf photosynthesis divided by transpiration. The g~1~ parameter was estimated from ITE to VPD response curves by fitting a rearranged optimal stomatal conductance model for ITE [@medlyn2012reconciling] using non-linear resgression, where K=0.5 [see @duursma2013near].

@@

## Leaf photosynthesis rates and photosynthetic capacity
Mean leaf photosynthesis rates were significantly higher in sun compared to shade leaves (+23%), under their local light environoment (Table 2). Following an increase in available light, A of shade leaves was significantly greater than both shade leaves at low light and sun leaves (P < 0.001). No effect of the warming treatment was detected on rates of A of sun leaves or shade leaves at low or high light. Photosynthesis within leaf types and treatments was similar through time and across the range of leaf temperatures measured (Figure `r get1fig`a).

\
The photsynthetic parameters J~max~ and Vc~max~ were both significantly higher in sun than shade leaves (Table 1), as evident from AC~i~ curves from each chamber (Figure `r acifig`a). No effect of the warming treatment was detected on either parameter (Table 1). Within individual chambers Vc~max~ was postively related to leaf N~area~ across leaf types and temperature treatments (P = 0.01, Figure `r acifig`b).  Additionally, leaf N~area~ was postively related to A across gas exchange campaigns and leaf types (P < 0.001, Figure `r acifig`c).  As Vc~max~ is condisdered implicit with g~m~ (results below) with conventional AC~i~ curves, a AC~c~ curves were also simulated to determine if treatment differences in AC~i~ parameters where instead the result of differences in g~m~.  When accounting for the effect of g~m~, photosynthetic capacity was still greater in sun leaves than shade leaves (Figure `r accfig`)

## Leaf water use efficiency
Leaf ITE was significantly greater in sun leaves than in shade leaves (P = 0.001, Table 2). Following an increase in available light, ITE of shade leaves did not differ from shade leaves at low light and was still significantly lower than sun leaves (P < 0.001).  ITE in sun leaves was reduced in the warming treatment (P = 0.021) but no effect was noted in shade leaves with low or high light. For all leaf types and light treatments there was a strong response of ITE to VPD, and individual data points broadly corresponded to response curves from the optimal ITE model (Figure `r waterusefig`a).  The mean estimated g~1~ for sun leaves was 1.51&pm;0.11 and for shade leaves with low and high light was 2.59&pm;0.12 and 2.74&pm;0.04.  Within leaf types and light treatments the response of VPD to leaf temperature was similar across all measurement campaigns (Figure `r get2fig`a)

\
Leaf $\delta$^13^C, as an index of water use efficiencey [@marshall2007sources], significantly decreased from sun leaves to shade leaves by ca. 2&permil; (Table 1). No effects of the warming treatment on leaf $\delta$^13^C were detected. Leaf $\delta$^13^C and N~area~ were postively correlated for all leaves (P<0.001, Figure `r waterusefig`b), with less negative $\delta$^13^C and higher N investment in sun leaves.

## Stomatal conductance, mesophyll conductance and CO~2~ drawdown
Mean g~s~ was significantly higher in shade compared to sun leaves (+18%), under their local light environment (P = 0.004, Table2).  Following an increase in available light, g~s~ of shade leaves was significantly greater than both shade leaves at low light and sun leaves (P < 0.001, Figure `r gebarfig`a). In sun leaves, g~s~ and A exhibited optimal behavior with diminishing rates of A with highest values of g~s~ (Figure `r agmgsfig`a).  However, in shade leaves this relationship was clearly uncoupled.  With increased light availability, both increases in A and g~s~ in shade leaves extened the relationship of A and g~s~ seen in sun leaves.

\
The observed carbon isotope discrimination was postively correlated with C~i~:C~a~ for all leaf types (P < 0.001) with larger $\Delta$~o~ detected for sun leaves and shade leaves at high light (Figure `r deltafig`). Here, the carbon isotope discrimination associtaed with g~m~ accounted for the majority of $\Delta$~o~ (69.7&pm;0.4 %) and varied little across measurement temperatures, leaf types, or warming treatments.

\
Mean g~m~ was higher in sun compared to shade leaves (+27%), under their local light environment ( Table2).  Following an increase in available light, g~s~ of shade leaves was significantly greater than both shade leaves at low light and sun leaves (P < 0.001). Relationships between g~m~ and A were nearly proportional within leaf types and light treatments (Figure `r gebarfig`b,c). Photosynthesis scaled postively with increases in g~m~ for all leaves and the large increases in g~m~ in shade leaves under high light resulted in the highest rates of A (Figure `r agmgsfig`b). No differences in g~s~ nor g~m~ were detected with the warming treatment within leaf types. Additionally, neither g~s~ nor g~m~ varied significantly across measurements campaigns (Figures `r get1fig`b and `r get2fig`b, respectively) and a slightly negative relationship with g~m~ and increasing leaf temperature was detected.

\
The combination of a non-limiting supply of leaf K and more open stomata lead to a significant increase in C~i~ in shade compared to sun leaves (Figure `r ciccfig`a). As increases in A in shade leaves were associated with higher rates of g~s~ the C~i~ was similar than under low light but still greater than sun leaves. As the drop in CO~2~ for intecellular spaces to the cholorplast, C~i~-C~c~, measures the coordiation between g~m~ and A [@caemmerer2014temperature] an increase in this drawdown was detected in shade leaves at high light (Figure `r ciccfig`b,c).  This was a direct result of the increase in g~m~ and its subsequent direct effect on A, when compared to both sun and shade leaves that are acclimated to their local light environment. CO~2~ drawdown from C~a~ to C~i~ and C~i~ to C~c~ were both relatively stable acorss the range of temperatures measured and gas exchange campaigns (Figure `r get2fig`c and `r get1fig`c, respectively).

@@

## sub heading
The allocation of N constrains photosynthesis rates and is thus a key trait in determining the relative contribution of individual leaves to canopy C gain.  Decreasing light availability should decrease the investment into photosynthetic enzyme within a canopy (Mooney and Gulmon 1979).  As a result, acclimation of photosynthetic capacity to irradiance is typically reflected in the key photosynthetic biochemical parameters Vc~max~ and J~max~ [@farquhar1980biochemical]. Our data agree with these conventional conclusions as the distribution of N, both biochemical parameters and A were all reduced in shade leaves.

@@

\
The pattern of wasteful water use in shade leaves is important as we hypothesized that g~s~ and A would be proportional across sun and shade leaves.  In sun leaves variation in A and g~s~ were strongly correlated, exhibiting behaviour agreeing with optimal stomatal theory. However, lower rates of A in shade leaves were not coupled with decreases in g~s~, explaining the observed decreases in WUE. This is significant as optimal stomatal behaviour has been reported across a wide range of ecosystems and plant functional types; however, empirical data is often collected only on sun leaves [see @prentice2014balancing; @lin2015optimal].  As a result, the often used economic framework of balancing costs of using water versus N allocation to predict A [@wright2003least] may break down when considering all leaves within a tree canopy.

@@

\
Increasing the drawdown of CO~2~ into the intercellular air spaces via g~s~, however, does not necessarily infer concomitant increases in A. This is because CO~2~ supply to the site of carboxylation, via g~m~, may be anatomically constrained. In this study A and g~m~ scaled positively across leaf types and increased proportionately with increased light intensity in shade leaves.  Research has suggested that aquaporins can facilitate increases in the CO~2~ permeability of the cell membranes resulting in rapid modulation of mesophyll conductance [@hanba2004overexpression; @heinen2009role; @li2014plant]. This provides a potential explanation for the observed rapid increases in g~m~ but is laregly untested in leaves of tree species. Our findings support a growing wealth of evidence that g~m~ is highly variable and can respond to environmental variables [@flexas2008mesophyll]. Here we provide empirical data showing g~m~ not only varies within a canopy but the up-regulation of g~m~ plays a critical role in the photosynthetic response of shade leaves to sunflecks.

\
Similar to Laisk [-@laisk2005adjustment] we provide evidence that optimal acclimation, in this case physiological adjustment of the drawdown of CO~2~ in shade leaves, is possibly directed towards occasional maximums of light availability over adaptations towards a low light environment.  If shade leaves "lie in wait" for sunflecks then perhaps we should consider an alternate leaf economic strategy to maximise C gain, beyond conventional trade-offs associated with canopy resource distribution.  If the valuation of sunflecks as a C resource is large enough then costs of sub-optimal stomatal behaviour could be offset over the leaf lifespan. For example, the potential benefits of C gain in leaves constrained by low light to levels equivalent to sun leaves may be large enough to tolerate excess losses from transpiration.  However, accounting for the heterogeneous nature of light within a canopy remains a current challenge for empirical and modelling studies. Thus, models which predict leaf photosynthesis from N distribution within a canopy will be flawed unless inclusion of canopy light extinction and the integrations of sunflecks on shade leaves are included [@de1997simple].

\
Furthermore, our findings have important implications for process based models that predict canopy C gain from rates of leaf photosynthesis. Although the distributions of resources, such as N and water, define leaf photosynthetic capacity it is the physiological behaviour of individual leaves which actually determine C gain. The observed physiological behaviour leaves in the shade, at both low and high irradiance, suggests that current theories of leaf optimal behaviour are incomplete.  Furthermore, g~m~ must now be considered as a dynamic process in modelling whole canopy C gain.  Additional empirical data are needed to determine both the mechanisms and the capacity of g~m~ to rapidly increase CO~2~ drawdown. To improve our ability to predict whole canopy C gain future research should prioritize the incorporation of both sun and shade leaf physiology, which may be optimized differently.

@@

**Table 1**. *Eucalyptus tereticornis* leaf traits and the distribution of resources between full sun and shade leaves under ambient and elevated temperature treatments. Leaf mass per area, leaf Nitrogen~area~, $\delta$^13^C, $\Psi$~pd~, $\Psi$~l~ and K values represent treatment mean (&pm; 1 standard error) across measurement campaigns (n=6). Values of Vc~max~ and J~max~ are treament mean(standard error) from AC~i~ curves measured in each chamber before the drought treatment. Different letters represent significant differences between treatments. The P value represents the overall difference between leaf types and warming treatments.

@@

\
**Figure 7**.  Relationship between the observed discrimination of ^13^CO~2~ and measured C~i~/C~a~ for sun leaves and shade leaves at both low and high light.  The solid line represents the the theortical line for C3 plants from [@evans1986carbon].

@@

\
**Figure S1**. Reponse of photosynthesis (a), mesophyll conductance (b) and C~i~-C~c~ to leaf temperature for sun leaves and shade leaves at low and high light.  Shaded symbols represents each monthyl measurement campaign.

\
**Figure S2**. Reponse of vapor pressure deficit (a), stomatal conductance (b) and C~a~-C~i~ to leaf temperature for sun leaves and shade leaves at low and high light.  Shaded symbols represents each monthyl measurement campaign.

\
**Figure S3**. Photosynthetic CO~2~ response (AC~c~) curves for sun and shade leaves at elevated and ambient temperature treatments. C~c~ values were predicited with mesophyll conductance, thus curves represent chloroplastic photosynthetic parameters at 25°C and at saturating light (1800 $\mu$mols m^-1^ s^-1^).

>>>>

---
title: "Response of mesophyll conductance to light availability allows shade leaves to lie in wait"
author: ""
date: ""
output:
  word_document:
    reference_docx: manuscriptstyle.docx
  pdf_document: default
  html_document:
    number_sections: yes
csl: new-phytologist.csl
bibliography: references.bib
---

@@

\
In this study we use *Eucalyptus tereticornis* Sm. trees, planted in climate controlled whole tree chambers with ambient and elevated temperature (+3&deg;C) treatments, to empirically evaluate the distribution of resources and leaf physiological behaviour of sun and shade leaves. Our hypotheses are as follows:

@@

\
Leaf level gas exchange was measured with a standard 6 cm^2^ leaf chamber using a portable gas exchange system (LI-6400, LI-COR, Lincoln, NE, USA).  This system was coupled with a tunable diode laser (TDL;TGA100,Campbell Scientific, Inc., Logan, UT, USA) for concurrent measurements of online C isotope discrimination. The CO~2~ in the leaf chamber was set at ambient atmospheric [CO~2~] (400 ppm) with a flow rate of 200 $\mu$mols s^-1^.  Two identical gas exchanges systems were run simultaneously, one in each of a randomly chosen WTC for each temperature treatment. Leaf temperatures were controlled at the current AT or ET chamber air temperature. PPFD in the chamber was set to match the individual light environment of each leaf type (explained above). Sun flecks were then simulated for shade leaves by increasing the chamber PPFD to the light environment of the full sun leaf in the same tree. The maximum sunfleck response was then recorded once shade leaves re-stabilized in the leaf cuvette (ca. 25 min).

@@

where $\Delta$~o~ is the observed discrimination and  $\Delta$~i~, $\Delta$~gm~ , $\Delta$~e~ and $\Delta$~f~ are the contributions to fractionation if C~i~ = C~c~, g~m~, respiration and photorespiration, respectively. The equations for each are as follows:

\
(5)
$$
\begin{aligned}
\Delta_i = &\frac{1}{1-t}a_i+\frac{1}{1-t}((1+t)b-a')\frac{C_i}{C_a}
\end{aligned}
$$

@@

where C~s~ is the CO~2~ partial pressure at the leaf surface, a~b~ (2.9&permil;) is the fractionation from boundary layer diffusion and a is the fractionation due to diffusion in air (4.4&permil;) [@evans1986carbon]. C~a~ and C~i~ are the atmospheric and intercellular partial pressures and $\Gamma_{*}$ is the compensation point in the absence of mitochondrial respiration in the light (R~d~). In this experiment both $\Gamma_{*}$ and R~d~ were derived using a standard Arrhenius function with parameters for *Eucalyptus globulis* from [@crous2012light]. The ternary effect corrections (t) are described by:

@@

\
The CO~2~ diffusion from the intercellular airspace to the chloroplast, g~m~, is given by its relationship to the CO~2~ assimilation rate (A) by:

@@

where C~c~ is the chloroplast CO~2~ partial pressure. Once g~m~ was calculated C~c~ and the drawdown of CO~2~ from the intercellular air spaces to the site of carboxylation were then estimated using Equation 11. Examples of this approach to measure gas exchange and C isotope discrimination are presented in [@evans2013temperature]. The variation in $\Delta$~o~ between sun and shade leaves and the simulated sunfleck where then compared as a function of C~i~:C~a~.

@@

where the term Hp~w~g represents the gravitational potential assumed to be minimal in this experiment due to small tree size [@whitehead1998regulation]. Leaf level instantaneous transpiration efficiency (ITE) was calculated as leaf photosynthesis divided by transpiration. The g~1~ parameter was estimated from ITE to VPD response curves by fitting a rearranged optimal stomatal conductance model for ITE [@medlyn2012reconciling] using non-linear regression, where K=0.5 [see @duursma2013near].

@@

## Leaf photosynthesis rates and photosynthetic capacity
Mean leaf photosynthesis rates were significantly higher in sun compared to shade leaves (+23%), under their local light environment (Table 2). Following an increase in available light, A of shade leaves was significantly greater than both shade leaves at low light and sun leaves (P < 0.001). No effect of the warming treatment was detected on rates of A of sun leaves or shade leaves at low or high light. Photosynthesis within leaf types and treatments was similar through time and across the range of leaf temperatures measured (Figure `r get1fig`a).

\
The photosynthetic parameters J~max~ and Vc~max~ were both significantly higher in sun than shade leaves (Table 1), as evident from AC~i~ curves from each chamber (Figure `r acifig`a). Within leaf types, no effect of the warming treatment was detected on either parameter. Within individual chambers Vc~max~ was positively related to leaf N~area~ across leaf types and temperature treatments (P = 0.01, Figure `r acifig`b).  Additionally, leaf N~area~ was positively related to A across gas exchange campaigns and leaf types (P < 0.001, Figure `r acifig`c).  As Vc~max~ is considered implicit with g~m~ (results below) with conventional AC~i~ curves, a AC~c~ curves were also simulated to determine if treatment differences in AC~i~ parameters where instead the result of differences in g~m~.  When accounting for the effect of g~m~, photosynthetic capacity was still greater in sun leaves than shade leaves (Figure `r accfig`)

## Leaf water use efficiency
Leaf ITE was significantly greater in sun leaves than in shade leaves at low light (P = 0.001). Following an increase in available light, ITE of shade leaves did not differ from shade leaves at low light and was still significantly lower than sun leaves (P < 0.001).  ITE in sun leaves was reduced in the warming treatment but no effect was noted in shade leaves with low or high light (Table 2). For all leaf types and light treatments there was a strong response of ITE to VPD, and individual data points broadly corresponded to response curves from the optimal ITE model (Figure `r waterusefig`a).  The mean estimated g~1~ for sun leaves was 1.51&pm;0.11 and for shade leaves with low and high light was 2.59&pm;0.12 and 2.74&pm;0.04.  Within leaf types and light treatments the response of VPD to leaf temperature was similar across all measurement campaigns (Figure `r get2fig`a)

\
Leaf $\delta$^13^C, as an index of water use efficiency [@marshall2007sources], significantly decreased from sun leaves to shade leaves by ca. 2&permil; (Table 1). No effects of the warming treatment on leaf $\delta$^13^C were detected. Leaf $\delta$^13^C and N~area~ were positively correlated for all leaves (P<0.001, Figure `r waterusefig`b), with less negative $\delta$^13^C and higher N investment in sun leaves.

## Leaf physiology
Mean g~s~ was significantly higher in shade compared to sun leaves (+18%) under their local light environment (P = 0.004). Increases in g~s~ led to increases in A in sun leaves, however, g~s~ and A were not correlated in shade leaves under low light (Figure `r agmgsfig`a). Following an increase in available light, g~s~ of shade leaves was significantly greater than both shade leaves at low light and sun leaves ( Figure `r gebarfig`a). With increased light availability, both increases in A and g~s~ in shade leaves exhibited similar behavior as observed in sun leaves.

\
The observed carbon isotope discrimination was positively correlated with C~i~:C~a~ for all leaf types (P < 0.001) with larger $\Delta$~o~ detected for sun leaves and shade leaves at high light (Figure `r deltafig`). Here, the carbon isotope discrimination associated with g~m~ accounted for the majority of $\Delta$~o~ (69.7&pm;0.4 %) and varied little across measurement temperatures, leaf types, or warming treatments.

\
Mean g~m~ was higher in sun compared to shade leaves (+27%) under their local light environment (P < 0.001).  Following an increase in available light, g~s~ of shade leaves was significantly greater than both shade leaves at low light and sun leaves (Table 2). Relationships between g~m~ and A were nearly proportional within leaf types and light treatments (Figure `r gebarfig`b,c). Photosynthesis scaled positively with increases in g~m~ for all leaves and the large increases in g~m~ in shade leaves under high light resulted in the highest rates of A (Figure `r agmgsfig`b). No differences in g~s~ nor g~m~ were detected with the warming treatment within leaf types. Additionally, neither g~s~ nor g~m~ varied significantly across measurements campaigns (Figures `r get1fig`b and `r get2fig`b, respectively) and a slightly negative relationship with g~m~ and increasing leaf temperature was detected.

\
The combination of a non-limiting supply of leaf K and more open stomata lead to a significant increase in C~i~ in shade compared to sun leaves (Figure `r ciccfig`a). As increases in A in shade leaves were associated with higher rates of g~s~ the C~i~ was similar than under low light but still greater than sun leaves. As the drop in CO~2~ for intecellular spaces to the cholorplast, C~i~-C~c~, measures the coordination between g~m~ and A [@caemmerer2014temperature] an increase in this drawdown was detected in shade leaves at high light (Figure `r ciccfig`b,c).  This was a direct result of the increase in g~m~ and its subsequent direct effect on A, when compared to both sun and shade leaves that are acclimated to their local light environment. CO~2~ drawdown from C~a~ to C~i~ and C~i~ to C~c~ were both relatively stable across the range of temperatures measured and gas exchange campaigns (Figure `r get2fig`c and `r get1fig`c, respectively).

@@

## Resource distribution and photosynthetic capacity
The allocation of N constrains photosynthesis rates and is thus a key trait in determining the relative contribution of individual leaves to canopy C gain.  Decreasing light availability should decrease the investment into photosynthetic enzyme within a canopy (Mooney and Gulmon 1979).  As a result, acclimation of photosynthetic capacity to irradiance is typically reflected in the key photosynthetic biochemical parameters Vc~max~ and J~max~ [@farquhar1980biochemical]. Our data agree with these conventional conclusions as the distribution of N, both biochemical parameters and A were all reduced in shade leaves.

@@

## Leaf physiology of sun and shade leaves
\
The pattern of wasteful water use in shade leaves is important as we hypothesized that g~s~ and A would be proportional across sun and shade leaves.  In sun leaves variation in A and g~s~ were strongly correlated, exhibiting behaviour agreeing with optimal stomatal theory. However, lower rates of A in shade leaves were not coupled with decreases in g~s~, explaining the observed decreases in WUE. This is significant as optimal stomatal behaviour has been reported across a wide range of ecosystems and plant functional types; however, empirical data is often collected only on sun leaves [see @prentice2014balancing; @lin2015optimal].  As a result, the often used economic framework of balancing costs of using water versus N allocation to predict A [@wright2003least] may break down when considering all leaves within a tree canopy.

@@

\
Increasing the drawdown of CO~2~ into the intercellular air spaces via g~s~, however, does not necessarily infer concomitant increases in A. This is because CO~2~ supply to the site of carboxylation, via g~m~, may be anatomically constrained. In this study A and g~m~ scaled positively across leaf types and increased proportionately with increased light intensity in shade leaves.  Research has suggested that aquaporins can facilitate increases in the CO~2~ permeability of the cell membranes resulting in rapid modulation of mesophyll conductance [@hanba2004overexpression; @heinen2009role; @li2014plant]. This provides a potential explanation for the observed rapid increases in g~m~ but is largely untested in leaves of tree species. Our findings support a growing wealth of evidence that g~m~ is highly variable and can respond to environmental variables [@flexas2008mesophyll]. Here we provide empirical data showing g~m~ not only varies within a canopy but the up-regulation of g~m~ plays a critical role in the photosynthetic response of shade leaves to sunflecks.

## Conclusions
\
Similar to Laisk [-@laisk2005adjustment] we provide evidence that optimal acclimation, in this case physiological adjustment of the drawdown of CO~2~ in shade leaves, is possibly directed towards occasional maximums of light availability over adaptations towards a low light environment.  If shade leaves "lie in wait" for sunflecks then perhaps we should consider an alternate leaf economic strategy to maximize C gain, beyond conventional trade-offs associated with canopy resource distribution.  If the valuation of sunflecks as a C resource is large enough then costs of sub-optimal stomatal behaviour could be offset over the leaf lifespan. For example, the potential C gain in leaves which experience high amounts of sunflecks may be large enough to tolerate excess losses from transpiration.  However, accounting for the heterogeneous nature of light within a canopy remains a current challenge for empirical and modelling studies. Thus, models which predict leaf photosynthesis from N distribution within a canopy will be flawed unless inclusion of canopy light extinction and the integration of sunflecks on shade leaves are included [@de1997simple].

\
Furthermore, our findings have important implications for process based models that predict canopy C gain from rates of leaf photosynthesis. Although the distributions of resources, such as N and water, define leaf photosynthetic capacity it is the physiological behaviour of individual leaves which actually determine C gain. The observed physiological behaviour leaves in the shade, at both low and high irradiance, suggests that current theories of leaf optimal behaviour are incomplete.  Furthermore, g~m~ must now be considered as a dynamic process in tree growth models that cannot be simply parameterized.  Additional empirical data are needed to determine both the mechanisms and the capacity of g~m~ to rapidly increase CO~2~ drawdown. To improve our ability to predict whole canopy C gain future research should prioritize the incorporation of both sun and shade leaf physiology, which may be optimized differently.

@@

**Table 1**. *Eucalyptus tereticornis* leaf traits and the distribution of resources between full sun and shade leaves under ambient and elevated temperature treatments. Leaf mass per area, leaf Nitrogen~area~, $\delta$^13^C, $\Psi$~pd~, $\Psi$~l~ and K values represent treatment mean (&pm; 1 standard error) across measurement campaigns (n=6). Values of Vc~max~ and J~max~ are treatment mean(standard error) from AC~i~ curves measured in each chamber at saturating light. Different letters represent significant differences between treatments. The P value represents the overall difference between leaf types and warming treatments.

@@

\
**Figure 7**.  Relationship between the observed discrimination of ^13^CO~2~ and measured C~i~/C~a~ for sun leaves and shade leaves at both low and high light.  The solid line represents the the theoretical line for C3 plants from [@evans1986carbon].

@@

\
**Figure S1**. Response of photosynthesis (a), mesophyll conductance (b) and C~i~-C~c~ to leaf temperature for sun leaves and shade leaves at low and high light.  Shaded symbols represents each monthly measurement campaign.

\
**Figure S2**. Response of vapor pressure deficit (a), stomatal conductance (b) and C~a~-C~i~ to leaf temperature for sun leaves and shade leaves at low and high light.  Shaded symbols represents each monthly measurement campaign.

\
**Figure S3**. Photosynthetic CO~2~ response (AC~c~) curves for sun and shade leaves at elevated and ambient temperature treatments. C~c~ values were predicted with mesophyll conductance, thus curves represent chloroplastic photosynthetic parameters at 25°C and at saturating light (1800 $\mu$mols m^-1^ s^-1^).
